### 2018-11-12 source/Ch0-Foreword/foreword.tex
asdasda

\pg
J.R.R Tolkien wrote, in his Mythopoeia, that ``He sees no stars who does not see them first, of living silver made that sudden burst, to flame like flowers beneath the ancient song"\footnote{See \href{http://home.agh.edu.pl/~evermind/jrrtolkien/mythopoeia.htm}{here} for the full text.}. In his defense of myth-making, he formulates the argument that the attribution of meaning is an act of creation - that ``trees are not `trees' until so named and seen" - and that this capacity for creation defines the human creature. The scientific endeavour, in this context, can be understood as a social expression of a fundamental feature of humanity, and from this endeavour flows much understanding. This thesis, one thread among many, focuses on the study of astronomical objects as seen by the radio waves they emit.

>>>>

\pg
J.R.R Tolkien wrote, in his Mythopoeia, that ``He sees no stars who does not see them first, of living silver made that sudden burst, to flame like flowers beneath the ancient song"\footnote{See \href{http://home.agh.edu.pl/~evermind/jrrtolkien/mythopoeia.htm}{here} for the full text: \url{http://home.agh.edu.pl/~evermind/jrrtolkien/mythopoeia.htm}.}. In his defense of myth-making, he formulates the argument that the attribution of meaning is an act of creation - that ``trees are not `trees' until so named and seen" - and that this capacity for creation defines the human creature. The scientific endeavour, in this context, can be understood as a social expression of a fundamental feature of humanity, and from this endeavour flows much understanding. This thesis, one thread among many, focuses on the study of astronomical objects as seen by the radio waves they emit.

### 2018-11-12 source/Ch1-ThisThesis/intro.tex
asdasda

\pg
Radio astronomy is a young but extremely dynamic field with a large range of science cases. From the study of the epoch of reionisation, which involves looking for polarised, redshifted 21cm line emission to study the emergence of the first luminous sources\footnote{See \url{https://www.haystack.mit.edu/ast/science/epoch/} for more information - the ultimate goal of this topic is to  learn more about the process of structure formation in the universe: what exactly are the evolutionary links between the very smooth matter distribution revealed by CMB studies and the highly structured universe of galaxies and clusters thereof which we know today?}, to that of transient radio sources such as pulsars and exoplanets, going through the study of ultra-high-energy cosmic rays, solar science, space weather science, the study of cosmic magnetism, and the creation of deep extragalactic surveys, radio-astronomy opens up an extremely fertile window into our Universe.

@@

\pg
So-called unified AGN models, which aim to explain the underlying physical object which is assumed to lie behind all observed examples of AGN, were developed to explain the diversity in observed AGN types \citepads[e.g.]{1985ApJ...297..621A,2008A&A...490..893T,2015ARA&A..53..365N,2017MNRAS.469.1883M}. These models would include a central engine, usually a supermassive black hole with hot accretion \citepads[innermost accretion regions estimated to be around $10^7$K by]{1972A&A....21....1P}, emitting thermally; an optically thick disk torus, typically of size 10 to 100pc, blocking light emitted from the central engine along the equitorial plane, and the cavity inside which would be the region from which broad emission lines would emit; and finally, ionisation cones in the direction of the poles, where the emission from the central engine would collimate into large-scale jets (which would also, at the scale of a few tens of parsecs, create the region of narrow spectral line emission).

\pg
The interaction between AGN jets and the surrounding intergalactic medium can result in structure such as that shown in \cref{fig.agn} below: a radio galaxy in the centre of two extremely radio-bright jets, with lots of diffuse emission from their lobes. These lobes show where the accretion-accelerated particles begin to slow down - and therefore emit - due to contact with the intergalactic medium. Hence, their sometimes peculiar shape: this medium can be complex and inhomogeneous (if the radio galaxies lie in a cluster, for example) which can lead to the sort of jet structure shown in this image.
\begin{figure}[!h]
\centering
\includegraphics[width=0.6\textwidth]{images/{3C31_radio_optical_montage_Alan_Bridle}.png}
\caption{\label{fig.agn} Radio optical montage of 3C31, by Alan Bridle: \url{https://www.cv.nrao.edu/~abridle/3c31xopt.htm}.}
\end{figure}

@@

\pg
The work done in this thesis can be summarised as the development of a new algorithm for mitigating calibration artefacts in radio-interferometric images, and its application to a peculiar famous extragalactic field. Because our data volumes put radio interferometry in the domain of ``Big Data" (there are $\sim$245TB of LOFAR data centred on the EGS, for example), there is a very real need for fast, reliable methods with which to improve interferometric images. This algorithm follows from an improved understanding of the noise properties of interferometric images, and can be understood as a way to better constrain the flux distribution in images by minimising the presence of artefacts. This, in turn, helps self-calibration convergence by improving the image resulting from each imaging step of the self-calibration process.

@@

\pg
The key finding of our algorithmic work is a fundamental relationship between the covariance matrices of residual visibilities and the covariance of pixels in the image-plane: the ``Cov-Cov relationship'' between visibility covariance and image-plane covariance. The pixel variance and covariance in the image-plane are determined by a ``noise-PSF'', convolved with every pixel in the sky. This also means that thermal noise in interferometric images is correlated between pixels, and that this correlation is given exactly by the PSF: in the limit of an empty, non-deconvolved sky, we therefore properly model the thermal noise properties of its interferometric images. This noise-PSF is the product of the Fourier transforms of the calibration gain covariance matrix with each cell mapped not from $uv$ space to $lm$ coordinates but rather between their respective covariance spaces - from a new {differential Fourier plane (henceforth ``$\left(\deltu\deltv\right)$''-plane) to the image-plane covariance space $\deltl\deltm$. This image-plane covariance space describes} the variance in each pixel and the covariance between pixels\footnote{{The noise-PSF also relates $\delta w$ to $\delta n$, as shown in our matrix formalism, but we do not explicitly reference this in the text for the sake of brevity, as visibility space is usually referred to as ``the UV-plane'' in literature, rather than ``the UVW-space''.}}. It describes the expected calibration artefacts and noise level around each source, does not vary as a function of direction, and is convolved with each source in the field to yield the final error map. Because all unwanted (in our case, unphysical) signal can be thought of as noise, we will refer to {the pixel variance map} as the ``noise-map''.

@@

\pg
The first step towards acquiring a deep, high-resolution image of the Extended Groth Strip therefore of creating a good, high-resolution model of 3C295 at LOFAR frequencies and for long baselines. Creating this model is one of the key results of this thesis. Starting from a high-resolution VLA model of 3C295, subbands selected across the LOFAR bandwidth are self-calibrated in iteratively greater numbers until a given noise threshold is reached. The flux scale is then boot-strapped based on the results of \citet{arse}. This ensures that 3C295 will be adequately subtracted from all corrected visibilities - including international baselines - and thus that images made using these visibilities will not be polluted by sidelobes from 3C295. 
The work done towards this goal is shown in \cref{section.3c295}. %, and its result - a high-resolution spectral model of 3C295 - is shown below.
%
%\pg
%\textcolor{red}{show key result here}

\pg
Before launching the EGS imaging run using all calibrated datasets, we measure the impact of direction-dependent effects on known sources in the field. Two direction-dependent effects are expected to dominate: decorrelation and differential gains. The first is modeled by the algorithms used in this thesis, whereas the second is a function of the gain variability as a function of distance from the calibrator. Eight calibrator sources, lying at various distances and directions from the calibrator source, are examined. They are summarised in \cref{table.LOBOS.sources}, which is reproduced below: note that number 6 is actually 3C295.

>>>>

\pg
Radio astronomy is a young but extremely dynamic field with a large range of science cases. From the study of the epoch of reionisation, which involves looking for polarised, redshifted 21cm line emission to study the emergence of the first luminous sources\footnote{See \url{https://www.haystack.mit.edu/ast/science/epoch/} for more information - the ultimate goal of this topic is to  learn more about the process of structure formation in the universe: what exactly are the evolutionary links between the very smooth matter distribution revealed by CMB studies and the highly structured universe of galaxies and clusters thereof which we know today?}, to that of variable radio sources such as pulsars and (theoretically) exoplanets, going through the study of ultra-high-energy cosmic rays, solar science, space weather science, the study of cosmic magnetism, and the creation of deep extragalactic surveys, radio-astronomy opens up an extremely fertile window into our Universe.

@@

\pg
So-called unified AGN models, which aim to explain the underlying physical object which is assumed to lie behind all observed examples of AGN, were developed to explain the diversity in observed AGN types \citepads[e.g.]{1985ApJ...297..621A,2008A&A...490..893T,2015ARA&A..53..365N,2017MNRAS.469.1883M}. These models would include a central engine, usually a supermassive black hole with hot accretion \citepads[innermost accretion regions estimated to be around $10^7$K by]{1972A&A....21....1P}, emitting thermally; an optically thick disk torus, typically of size 10 to 100pc, blocking light emitted from the central engine along the equatorial plane, and the cavity inside which would be the region from which broad emission lines would emit; and finally, ionisation cones in the direction of the poles, where the emission from the central engine would collimate into large-scale jets (which would also, at the scale of a few tens of parsecs, create the region of narrow spectral line emission).

\pg
The interaction between AGN jets and the surrounding intergalactic medium can result in structure such as that shown in \cref{fig.agn} below: the bright nucleus of a radio galaxy in the centre of two extremely radio-bright jets, with lots of diffuse emission from their lobes. These lobes show where the accretion-accelerated particles begin to slow down - and therefore emit - due to contact with the intergalactic medium. Hence, their sometimes peculiar shape: this medium can be complex and inhomogeneous (if the radio galaxies lie in a cluster, for example) which can lead to the sort of jet structure shown in this image.
\begin{figure}[!h]
\centering
\includegraphics[width=0.6\textwidth]{images/{3C31_radio_optical_montage_Alan_Bridle}.png}
\caption{\label{fig.agn} Radio optical montage of 3C31, by Alan Bridle: \url{https://www.cv.nrao.edu/~abridle/3c31xopt.htm}.}
\end{figure}

@@

\pg
The work done in this thesis can be summarised as the development of a new algorithm for mitigating calibration artefacts in radio-interferometric images, and its application to a peculiar famous extragalactic field. Because our data volumes put radio interferometry in the domain of ``Big Data" (there are $\sim$245TB of LOFAR data centred on the Extended Groth Strip, for example), there is a very real need for fast, reliable methods with which to improve interferometric images. This algorithm follows from an improved understanding of the noise properties of interferometric images, and can be understood as a way to better constrain the flux distribution in images by minimising the presence of artefacts. This, in turn, helps self-calibration convergence by improving the image resulting from each imaging step of the self-calibration process.

@@

\pg
The key finding of our algorithmic work is a fundamental relationship between the covariance matrices of residual visibilities and the covariance of pixels in the image-plane: the ``Cov-Cov relationship'' between visibility covariance and image-plane covariance. The pixel variance and covariance in the image-plane are determined by a ``noise-PSF'', convolved with every pixel in the sky. This also means that thermal noise in interferometric images is correlated between pixels, and that this correlation is given exactly by the PSF: in the limit of an empty, non-deconvolved sky, we therefore properly model the thermal noise properties of its interferometric images. This noise-PSF is the product of the Fourier transforms of the calibration gain covariance matrix with each cell mapped not from $uv$ space to $lm$ coordinates\footnote{$uv$ space is where interferometric measurement coordinates lie, and $lm$ coordinates are the cardinal sines on which the inverse Fourier transform of the visibilities are mapped.} but rather between their respective covariance spaces - from a new {differential Fourier plane (henceforth ``$\left(\deltu\deltv\right)$''-plane) to the image-plane covariance space $\deltl\deltm$. This image-plane covariance space describes} the variance in each pixel and the covariance between pixels\footnote{{The noise-PSF also relates $\delta w$ to $\delta n$, as shown in our matrix formalism, but we do not explicitly reference this in the text for the sake of brevity, as visibility space is usually referred to as ``the UV-plane'' in literature, rather than ``the UVW-space''. For more information, see \cref{chapter.paper}}}. It describes the expected calibration artefacts and noise level around each source, does not vary as a function of direction, and is convolved with each source in the field to yield the final error map. Because all unwanted (in our case, unphysical) signal can be thought of as noise, we will refer to {the pixel variance map} as the ``noise-map''.

@@

\pg
The first step towards acquiring a deep, high-resolution image of the Extended Groth Strip is therefore to create a good, high-resolution model of 3C295 at LOFAR frequencies and for long baselines. Creating this model is one of the key results of this thesis. Starting from a high-resolution VLA model of 3C295, subbands selected across the LOFAR bandwidth are self-calibrated in iteratively greater numbers until a given noise threshold is reached. The flux scale is then boot-strapped based on the results of \citet{arse}. This ensures that 3C295 will be adequately subtracted from all corrected visibilities - including international baselines - and thus that images made using these visibilities will not be polluted by sidelobes from 3C295. 
The work done towards this goal is shown in \cref{section.3c295}. %, and its result - a high-resolution spectral model of 3C295 - is shown below.
%
%\pg
%\textcolor{red}{show key result here}

\pg
Before launching the EGS imaging run using all calibrated datasets, we measure the impact of direction-dependent effects on known sources in the field. Two direction-dependent effects are expected to dominate: decorrelation and differential gains. The first is modeled by the algorithms used in this thesis, whereas the second is a function of the gain variability as a function of distance from the calibrator. Eight calibrator sources, lying at various distances and directions from the calibrator source, are examined. They are summarised in \cref{table.LOBOS.sources}, which is reproduced below: note that number 6 is actually 3C295. Their distribution in the sky is shown in \cref{lbcs-coverage-image11}.

### 2018-11-12 source/Ch2-techniques/RIME.tex
asdasda

\pg
In this section, we introduce the mathematical framework which forms the basis for our algorithmic work.
While radio interferometry has historically been a complex business \citepads{2001isra.book.....T}, modern radio interferometry is based on the Radio Interferometer's Measurement Equation, a powerful and elegant formalism which underpins modern calibration and imaging algorithms. In this section, I will give a simplified account of this formalism as described in \citetads{2011A&A...527A.106S} and its companion papers\footnote{\citetads{2011A&A...527A.107S}, \citetads{2011A&A...527A.108S} and \citetads{2011A&A...531A.159S}}. This will form the theoretical basis on which further sections will expand. I cannot overstate the importance of the RIME as the foundation of modern calibration and imaging algorithms.

@@

\pg
Consider a single quasi-monochromatic point source in the sky. Its signal at a point in time and space can be described by a complex 2-vector $\evector$ -- this assumes plane waves. We can then represent physical processes which affect this signal's propagation using \emph{Jones matrices}. In other words:
\begin{equation}
\volt = \Jones\evector
\end{equation}
where $\volt$ is the voltage measured by our antenna, and $\Jones$ the Jones matrix describing the \emph{net} propagation effects - atmospheric, astrophysical, and instrumental - which affect our source's signal. $\Jones$ can be written as a series of matrix products, each individual matrix describing a physical phenomenon. They are then called a \emph{Jones chain}, and the resulting $\Jones$ is often called the \emph{total Jones matrix}.

\pg
A visibility is the correlation of voltage from two antennas. In other words,
\begin{align}
\Vis_{pq} &= 2 \langle\volt_p (\volt_q)^\Herm\rangle\\
\Vis_{pq} &= 2 \langle\Jones_p \evector (\Jones_q \evector)^\Herm\rangle\\
\Vis_{pq} &= 2 \langle\Jones_p \evector \evector^\Herm \Jones_q^\Herm\rangle
\end{align}
where $\Jones_p$ corresponds to the signal propagation chain of antenna $p$, and $\Jones_q$ for antenna $q$. Note that a factor of 2 has been introduced here, as a matter of convention. This is due to the definition of the Stokes parameter in relation to our $\evector$ outer product.

\pg
For $\mathbf{e}$ to be a 2-vector, rather than a 3-vector, we must choose to use $xyz$ as our coordinates, with $z$ the direction of propagation of our signal. \citetads{1996A&AS..117..137H} then show that the following relation holds:
\begin{equation}
2 \evector \evector^\Herm = 2  \begin{bmatrix} <e_x e_x^\star> & <e_x e_y^\star> \\ <e_y e_x^\star> & <e_y e_y^\star> \end{bmatrix} = \begin{bmatrix} I+Q & U+iV \\ U-iV & I-Q \end{bmatrix} = \Bmatrix
\end{equation}

@@

\section{Expanding our Jones Chain}
\label{section.RIME.JonesChain}
\pg
Having written our RIME as a matrix multiplication problem, we can now begin to differentiate between different types of Jones matrices. $\Jones_p$ describes the \emph{aggregate} effects which occur over the course of the signal's propagation to our antenna $p$. It can be split into a Jones chain of $n$ separate matrices, corresponding to different effects:
\begin{align}
\displaystyle \Jones_{sp}^H &= \prod_1^n \Jones_{nsp}^H = \Jones_{1sp}^H \Jones_{2sp}^H ... \Jones_{nsp}^H\\
\displaystyle \Jones_{sp}   &= \Jones_{nsp}...\Jones_{2sp} \Jones_{1sp}
\end{align}
where $s$ is a particular direction in the sky: these Jones matrices are therefore direction-dependent.
Since the Jones terms are applied sequentially to the brightness matrix, starting with $\Jones_{1sp}$, the terms to the right of our decomposition of $\Jones_{sp}$ can be said to occur `at the source', i.e. before those to the left (said to occur `at the antenna').

@@

\pg
Here, we have decoupled the nominal phase delay term ($\Kjones_p$) from all `corrupting' effects ($\Gjones_p$). We can take this one step further by defining the \emph{source coherency}, $\Coherency_{pq}$, as follows:
\begin{align}\label{eq.def.coherency}
\Coherency_{pq} &= \Kjones_p \Bmatrix_{pq} \Kjones_p^H\\
\Vis_{pq} &= \Gjones_p \Coherency_{pq} \Gjones_q^H
\end{align}

@@

\pg
We now expand our Jones chain as before. Some, but not all, elements in the chain will only depend on the antenna. These will be the same for all sources. We can thus write our Jones chain as:
\begin{equation}
\Jones_{sp} = \Gjones_p \Ejones_{sp} \Kjones_{sp}
\end{equation}

\pg
Note that we have kept the source-dependent effects (i.e. $E_{sp}$ those occurring `at the source') are to the right of our chain, and the antenna-dependent ($G_p$) effects to the left. $K_{sp}$ is a scalar and can thus be moved at will; it is therefore placed at the far right to recover the coherency matrix if necessary.

@@

\pg
This `onion' form of the RIME is usually sufficient to describe signal propagation in radio-interferometric arrays. We find the three types of Jones matrices promised earlier: the geometric effects accounted for by $\Kjones$, the direction-dependent effects modeled into $\Ejones$, and the direction-independent effects (antenna-based effects) modeled into $\Gjones$.

@@

\pg
Having written this, and using $\lvec$ as shorthand $(l,m)$, we can once again decompose our Jones matrix into a Jones chain containing a direction-independent term $\Gjones$, a direction-dependent term $\Ejones$, and the phase term $\Kjones$. Substituting this into our integral, we get the following expression:
\begin{align}
\Vis_{pq} &= \Gjones_p \bigg(\int_{\lvec} \frac{1}{n} \Ejones_p(\lvec) \Bmatrix \Ejones_q(\lvec) \Kjones_p(\lvec)\Kjones_q^H(\lvec) dl dm \bigg) \Gjones_q^H\\
          &= \Gjones_p \bigg(\int_{\lvec} \frac{1}{n} \Ejones_p(\lvec) \Bmatrix \Ejones_q(\lvec) e^{-2\pi i\big(u_{pq}l+v_{pq}m+w_{pq}(n-1)\big)} dl dm \bigg) \Gjones_q^H
\end{align}
where
\begin{alignat}{3}
u_{pq}&=u_p-u_q \qquad v_{pq}&=v_p-v_q \qquad w_{pq}&=w_p-w_q
\end{alignat}

@@

\pg
Earlier calibration models and algorithms rely on the hypothesis that all baselines see the same sky. This is the \emph{fundamental premise of self-calibration}. The topic of self-cal is covered in \cref{section.calibration.2gc}. This premise only holds when all Direction-Dependent Effects (henceforth referred to as DDEs) are identical for all baselines (and therefore for all antennas). This is a necessary condition for the apparent sky $\Bmatrix_{pq}$ to be the same for all baselines\footnote{$\Bmatrix_{pq}$ would then correspond to the `true' sky attenuated by the ``power beam", which corresponds to our instrument's directional sensitivity in the traditional view}, as it allows us to write:
\begin{align}
\Ejones_p(\lvec) \Wjones_p(\lvec)&= \Ejones(\lvec) \Wjones(\lvec)\\
\Bmatrix_{pq}(\lvec) = \Bmatrix_{app}(\lvec) &= \Ejones(\lvec) \Wjones(\lvec) \Bmatrix)(\lvec) \Wjones^H(\lvec) \Ejones^H(\lvec)
\end{align}

\pg
If this condition is met, we can then rewrite the full-sky RIME as:
\begin{equation}\label{eq.fullsky.RIME}
\Vis_{pq} = \Gjones_p \Coherency_{pq} \Gjones_q^H
\end{equation}
where $\Coherency_{pq} = \Coherency(u_{pq},v_{pq}) = \Coherency(\uvec)$. Comparing \cref{eq.def.coherency,eq.vis.FT.coherency.matrix} directly shows that the coherency matrix is simply the two-dimensional Fourier transform of $\Bmatrix_{app}(\lvec)$. We can thus call $\Coherency(\uvec)$ the \emph{sky coherency}, in keeping with our nomenclature for the RIME of a point source.

@@

\pg
$\Jones(t,\nu)$ describes the propagation of an electromagnetic signal through various physical processes. This signal has a complex phase, which is variable in both time and frequency. Since it is a rotating vector, integrating over any time or frequency band will always result in a \emph{net loss} in amplitude in the measured $\langle\Vis\rangle$. This mechanism is commonly referred to as \emph{time/bandwidth width decorrelation} or \emph{smearing} \citepads{2011A&A...527A.106S}.

@@

\pg
Assuming that $\Delta t$ and $\Delta \nu$ are small enough that the amplitude of $\Vis_{pq}$ remains constant over the integration band (i.e. that $\Vis_{pq}$ is well-sampled by our instrument), then a useful first-order approximation is:
\begin{equation}
\langle \Vis_{pw} \rangle \simeq \text{sinc}\frac{\Delta \pmb{\Psi}}{2} \enskip \text{sinc}\frac{\Delta \pmb{\Phi}}{2} \enskip \Vis_{pq}(t_{mid},\nu_{mid})
\end{equation}
where
\begin{align}
t_{mid}           &= (t_0 + t_1)/2\\
\nu_{mid}         &= (\nu_0 + \nu_1)/2
\end{align}
\begin{align}
\Delta \pmb{\Psi} &= \arg \Vis_{pq}(t_1,\nu_{mid}) - \arg \Vis_{pq}(t_0, \nu_{mid})\\
\Delta \pmb{\Phi} &= \arg \Vis_{pq}(t_{mid},\nu_1) - \arg \Vis_{pq}(t_{mid}, \nu_0)
\end{align}

>>>>

\pg
In this section, we introduce the mathematical framework which forms the basis for our algorithmic work.
While radio interferometry has historically been a complex business \citepads{2001isra.book.....T}, modern radio interferometry is based on the Radio Interferometer's Measurement Equation, a powerful and elegant formalism which underpins modern calibration and imaging algorithms. In this section, I will give a simplified account of this formalism as described in \citetads{2011A&A...527A.106S} and its companion papers \citepads{2011A&A...527A.107S, 2011A&A...527A.108S, 2011A&A...531A.159S}. This will form the theoretical basis on which further sections will expand.% I cannot overstate the importance of the RIME as the foundation of modern calibration and imaging algorithms.

@@

\pg
Consider a single quasi-monochromatic point source in the sky. Its signal at a point in time and space can be described by a complex 2-vector $\evector$ -- this assumes plane waves. We can then represent physical processes which affect this signal's propagation using \emph{Jones matrices}. In other words:
\begin{align}
\volt &= \Jones\evector\\
\evector &= \begin{pmatrix} e_x \\ e_y \\ e_z \end{pmatrix}
\end{align}
where $\volt$ is the voltage measured by our antenna, and $\Jones$ the Jones matrix describing the \emph{net} propagation effects - atmospheric, astrophysical, and instrumental - which affect our source's signal. $\evector$ is the electromagnetic wave emitted by a source. $\Jones$ can be written as a series of matrix products, each individual matrix describing a physical phenomenon. They are then called a \emph{Jones chain}, and the resulting $\Jones$ is often called the \emph{total Jones matrix}.

\pg
For $\mathbf{e}$ to be a 2-vector, rather than a 3-vector, we must choose to use $xyz$ as our coordinates, with $z$ the direction of propagation of our signal. \citetads{1996A&AS..117..137H} then show that the following relation holds:
\begin{equation}
2 \evector \evector^\Herm = 2  \begin{bmatrix} <e_x e_x^\star> & <e_x e_y^\star> \\ <e_y e_x^\star> & <e_y e_y^\star> \end{bmatrix} = \begin{bmatrix} I+Q & U+iV \\ U-iV & I-Q \end{bmatrix} = \Bmatrix
\end{equation}
where $\Bmatrix$ is the coherency matrix of the source's emission.

\pg
A visibility is the correlation of voltage from two antennas. In other words,
\begin{align}
\Vis_{pq} &= 2 \langle\volt_p (\volt_q)^\Herm\rangle\\
\Vis_{pq} &= 2 \langle\Jones_p \evector (\Jones_q \evector)^\Herm\rangle\\
\Vis_{pq} &= 2 \langle\Jones_p \evector \evector^\Herm \Jones_q^\Herm\rangle
\end{align}
where $\Jones_p$ corresponds to the signal propagation chain of antenna $p$, and $\Jones_q$ for antenna $q$. $(...)^\Herm$ corresponds to taking the Hermitian conjugate of a matrix, and $\langle...\rangle$ an average over some interval. Note that a factor of 2 has been introduced here, as a matter of convention. This is due to the definition of the Stokes parameter in relation to our $\evector$ outer product.

@@

\section{Expanding our Jones Chain}
\label{section.RIME.JonesChain}
\pg
Having written our RIME as a matrix multiplication problem, we can now begin to differentiate between different types of Jones matrices. $\Jones_p$ describes the \emph{aggregate} effects which occur over the course of the signal's propagation to our antenna $p$. It can be split into a Jones chain of $n$ separate matrices, corresponding to different effects:
\begin{align}
\displaystyle \Jones_{sp}^H &= \prod_1^n \Jones_{nsp}^H = \Jones_{1sp}^H \Jones_{2sp}^H ... \Jones_{nsp}^H\label{eatshit1}\\
\displaystyle \Jones_{sp}   &= \Jones_{nsp}...\Jones_{2sp} \Jones_{1sp}\label{eatshit2}
\end{align}
where $s$ is a particular direction in the sky: these Jones matrices are therefore direction-dependent. Going from \cref{eatshit1} to \cref{eatshit2} simply consists of applying the Hermitian operator to both sides of \cref{eatshit1} to recover $\Jones_{nsp}$.
Since the Jones terms are applied sequentially to the brightness matrix, starting with $\Jones_{1sp}$, the terms to the right of our decomposition of $\Jones_{sp}$ can be said to occur `at the source', i.e. before those to the left (said to occur `at the antenna').

@@

\pg
Here, we have decoupled the nominal phase delay term ($\Kjones_p$) from all `corrupting' effects ($\Gjones_p$). We can take this one step further by defining the \emph{source coherency}, $\Coherency_{pq}$, as follows:
\begin{align}\label{eq.def.coherency}
\Coherency_{pq} &= \Kjones_p \Bmatrix \Kjones_p^H\\
\Vis_{pq} &= \Gjones_p \Coherency_{pq} \Gjones_q^H
\end{align}

@@

\pg
We now expand our Jones chain as before. Some, but not all, elements in the chain will only depend on the antenna. These will be the same for all sources. We can thus write our Jones chain as:
\begin{equation}
\Jones_{sp} = \Gjones_p \Ejones_{sp} \Kjones_{sp}
\end{equation}
where $E_{sp}$ are direction-dependent effects, which vary for each source. Note that we have kept the source-dependent effects (i.e. those which vary depending on the line of sight) are to the right of our chain, and the antenna-dependent ($G_p$) effects to the left. $K_{sp}$ is a scalar and can thus be moved at will; it is therefore placed at the far right to recover the coherency matrix if necessary.

@@

\pg
This `onion' form of the RIME - so-called because, like an onion, it is formed in layers - is usually sufficient to describe signal propagation in radio-interferometric arrays. We find the three types of Jones matrices promised earlier: the geometric effects accounted for by $\Kjones$, the direction-dependent effects modeled into $\Ejones$, and the direction-independent effects (antenna-based effects) modeled into $\Gjones$.

@@

\pg
Having written this, and using $\lvec$ as shorthand $(l,m)$, we can once again decompose our Jones matrix into a Jones chain containing a direction-independent term $\Gjones$, a direction-dependent term $\Ejones$, and the phase term $\Kjones$. Substituting this into our integral, we get the following expression:
\begin{align}
\Vis_{pq} &= \Gjones_p \bigg(\int_{\lvec} \frac{1}{n} \Ejones_p(\lvec) \Kjones_p(\lvec) \Bmatrix \Kjones_q^H(\lvec) \Ejones_q(\lvec)  dl dm \bigg) \Gjones_q^H\\\Vis_{pq} &= \Gjones_p \bigg(\int_{\lvec} \frac{1}{n} \Ejones_p(\lvec) \Bmatrix \Ejones_q(\lvec) \Kjones_p(\lvec)\Kjones_q^H(\lvec) dl dm \bigg) \Gjones_q^H\\
          &= \Gjones_p \bigg(\int_{\lvec} \frac{1}{n} \Ejones_p(\lvec) \Bmatrix \Ejones_q(\lvec) e^{-2\pi i\big(u_{pq}l+v_{pq}m+w_{pq}(n-1)\big)} dl dm \bigg) \Gjones_q^H
\end{align}
where
\begin{alignat}{3}
u_{pq}&=u_p-u_q \qquad v_{pq}&=v_p-v_q \qquad w_{pq}&=w_p-w_q
\end{alignat}

@@

\pg
Earlier calibration models and algorithms rely on the hypothesis that all baselines see the same sky. This is the \emph{fundamental premise of self-calibration}. The topic of self-cal is covered in \cref{section.calibration.2gc}. This premise only holds when all Direction-Dependent Effects (henceforth referred to as DDEs) are identical for all baselines (and therefore for all antennas). We then effectively have $\Ejones_{s,p}=1$ for all $p$. This is a necessary condition for the \textit{apparent} sky $\Bmatrix_{pq}$ to be the same for all baselines\footnote{The ``apparent" sky is the `true' sky attenuated by the ``power beam", which corresponds to our instrument's directional sensitivity in the traditional view.}, as it allows us to write:
\begin{align}
\Ejones_p(\lvec) \Wjones_p(\lvec)&= \Ejones(\lvec) \Wjones(\lvec)\\
\Bmatrix_{pq}(\lvec) = \Bmatrix_{app}(\lvec) &= \Ejones(\lvec) \Wjones(\lvec) \Bmatrix (\lvec) \Wjones^H(\lvec) \Ejones^H(\lvec)
\end{align}

\pg
If this condition is met, we can then rewrite the full-sky RIME as:
\begin{equation}\label{eq.fullsky.RIME}
\Vis_{pq} = \Gjones_p \Coherency_{pq} \Gjones_q^H
\end{equation}
where $\Coherency_{pq} = \Coherency(u_{pq},v_{pq}) = \Coherency(\uvec)$. Our coherency matrix includes the beam attenuation implicitly, as $\Bmatrix_{app}$ is now the coherency of the \textit{apparent} sky brightness distribution. Comparing \cref{eq.def.coherency,eq.vis.FT.coherency.matrix} directly shows that the coherency matrix is simply the two-dimensional Fourier transform of $\Bmatrix_{app}(\lvec)$. We can thus call $\Coherency(\uvec)$ the \emph{sky coherency}, in keeping with our nomenclature for the RIME of a point source.

@@

\pg
$\Jones(t,\nu)$ describes the propagation of an electromagnetic signal through various physical processes. This signal has a complex phase, which is variable in both time and frequency. Since it is a rotating vector, integrating over any time or frequency band will always result in a \emph{net loss} in amplitude in the measured $\langle\Vis\rangle$. This mechanism is commonly referred to as \emph{time/bandwidth decorrelation} or \emph{smearing} \citepads{2011A&A...527A.106S}.

@@

\pg
Assuming that $\Delta t$ and $\Delta \nu$ are small enough that the amplitude of $\Vis_{pq}$ remains constant over the integration band (i.e. that $\Vis_{pq}$ is well-sampled by our instrument), then a useful first-order approximation\footnote{see \citetads{2011A&A...527A.107S} and companion papers.} is:
\begin{equation}
\langle \Vis_{pw} \rangle \simeq \text{sinc}\frac{\Delta \pmb{\Psi}}{2} \enskip \text{sinc}\frac{\Delta \pmb{\Phi}}{2} \enskip \Vis_{pq}(t_{mid},\nu_{mid})
\end{equation}
where
\begin{align}
t_{mid}           &= (t_0 + t_1)/2\\
\nu_{mid}         &= (\nu_0 + \nu_1)/2
\end{align}
\begin{align}
\Delta \pmb{\Psi} &= \arg \Vis_{pq}(t_1,\nu_{mid}) - \arg \Vis_{pq}(t_0, \nu_{mid})\\
\Delta \pmb{\Phi} &= \arg \Vis_{pq}(t_{mid},\nu_1) - \arg \Vis_{pq}(t_{mid}, \nu_0)
\end{align}
